---
title: "Statistical Inference Course Project"
author: "gspinnadauhu"
date: "March 23, 2017"
output: pdf_document
---

#### Part 1: Simulation Exercise
1. Overview
In this project we are investigating the exponential distribution in R and compare it with the Central Limit Theorem. We are investigating distribution of averages for 40 expenonentials.
2. Simulations
Taking 1000 averages of 40 random exponentials with lambda=0.2. Putting averages in dataframe mns.
```{r}
mns<-NULL
for (i in 1:1000) {
        mns=c(mns,mean(rexp(40,0.2)))
}
mns<-data.frame(mns)
```
3. Sample Mean vs Theoretical Mean
The theoretical mean of an exponential distribution is 1/lambda (lambda=0.2 in this case).
Comparing the theoretical mean with the mean of sampled means.
```{r}
library(ggplot2)
ggplot(data=mns, aes(mns))+
        geom_histogram()+
        labs(title="Theoretical Mean vs Sample Mean",
             x="Sample Means",
             y="Count")+
        geom_vline(xintercept=1/0.2, colour="green",linetype="longdash", size=1)+
        geom_text(aes(x=5.5,y=10,label="Theoretical Mean"),color="green")+
        theme_classic()
```
4. Sample Variance vs Theoretical Variance
The theoretical variance of an exponential distribution is (1/lambda)^2. Comparing the theoretical variance with the variance of the sample means.
```{r}

```
5. Distribution
Comparing 1000 draws from an exponential distrion with lambda=0.2 with the distribution of the sampled means in mns.
```{r}

```

#### Part 2: Basic Inferential Data Analysis
1. Loading Data & Basic Analysis
2. Data Summary
3. Comparing Tooth Growth by supp & dose
4. Assumptions & Conclusions