---
title: "Statistical Inference Course Project"
author: "gspinnadauhu"
date: "March 23, 2017"
output: pdf_document
---

#### Part 1: Simulation Exercise
1. Overview
    In this project we are investigating the exponential distribution in R and compare it with the Central Limit Theorem. We are investigating distribution of averages for 40 expenonentials.
2. Simulations
    Taking 1000 averages and variances of 40 random exponential distributions with lambda=0.2. Putting samples in dataframe sims.
    ```{r}
    lambda<-0.2
    mns<-NULL
    vars<-NULL
    for (i in 1:1000) {
            mns=c(mns,mean(rexp(40,lambda)))
            vars=c(vars,var(rexp(40,lambda)))
    }
    sims<-data.frame(mns,vars)
    ```
3. Sample Mean vs Theoretical Mean
    The theoretical mean of an exponential distribution is 1/lambda (lambda=0.2 in this case).
    Comparing the theoretical mean with the mean of sampled means.
    ```{r}
    library(ggplot2)
    ggplot(data=sims, aes(mns))+
            geom_histogram()+
            labs(title="Theoretical Mean vs Sample Mean",
                 x="Sample Means",
                 y="Count")+
            geom_vline(xintercept=1/lambda, colour="green",linetype="longdash", size=1)+
            geom_text(aes(x=1/lambda*1.15,y=10,label="Theoretical Mean"),color="green")+
            theme_classic()
    ```
    As visible above, the sample means appear to center around the distribution's theoretical mean.
    Furthermore, the mean of the sample means is almost exactly equal to 1/lambda=1/0.2=5
    ```{r}
    mean(sims$mns)
    ```
4. Sample Variance vs Theoretical Variance
    The theoretical variance of an exponential distribution is (1/lambda)^2. Comparing the theoretical variance with the variance of the sample means.
    ```{r}
    library(ggplot2)
    ggplot(data=sims, aes(vars))+
            geom_histogram()+
            labs(title="Theoretical Variance vs Sample Variance",
                 x="Sample Variances",
                 y="Count")+
            geom_vline(xintercept=(1/lambda)^2, colour="blue",linetype="longdash", size=1)+
            geom_text(aes(x=(1/lambda)^2*1.6,y=10,label="Theoretical Variance"),color="blue")+
            theme_classic()
    ```
    As visible above, the sample variances appear are for the most part located around the theoretical variance.
    Furthermore, the mean of the sample variances is very close to (1/lambda)^2=(1/0.2)^2=25
    ```{r}
    mean(sims$vars)
    ```


5. Distribution
    Comparing 1000 draws from an exponential distribution with lambda=0.2 with the distribution of the 1000 sampled means of 40 exponentials in mns.
    First, simulating the 1000 draws, converting to a data frame & plotting the distribution.
    ```{r}
    draws<-rexp(1000,0.2)
    draws<-data.frame(draws)
    drawmean<-mean(draws$draws)
    library(ggplot2)
    ggplot(data=draws, aes(draws))+
            geom_histogram()+
            labs(title="1000 Random Draws from Exponential Distribution",
                 x="Samples",
                 y="Count")+
            geom_vline(xintercept=mean(drawmean), colour="lightblue",linetype="longdash", size=1)+
            geom_text(aes(x=drawmean+1.5,y=10,label="Mean"),color="lightblue")+
            theme_classic()
    ```
    These random draws conform to the expected shape of an exponential distribution.
    Now, comparing this to the distrbution of means drawn from 1000 simulated sets of 40 exponential distributions in sims$mns.
    ```{r}
    library(ggplot2)
    ggplot(data=sims, aes(mns))+
            geom_histogram()+
            labs(title="Distribution and Mean of 1000 Sampled Means",
                 x="Sample Means",
                 y="Count")+
            geom_vline(xintercept=1/lambda, colour="green",linetype="longdash", size=1)+
            geom_text(aes(x=mean(sims$mns)+.25,y=10,label="Mean"),color="green")+
            theme_classic()
    ```
    While both have a mean of 5, which matches the theoretical mean, the draws are normally distributed only in the set of 1000 means drawn for samples size 40. As the sample size increases, the sampled means conform closer to the theoretical mean, which is why the distribution looks more bell curved.

#### Part 2: Basic Inferential Data Analysis
This analysis is using the R ToothGrowth dataset which was created from assessing the effect of vitamin C on tooth growth in guinea pigs.
The dataset contains 60 observations with following variables:

* len - tooth length
* supp - supplement type (VC-ascorbic acid, OJ-orange juice)
* dose - dose in miligrams per day
    
1. Loading Data & Basic Analysis
    First, loading the ToothGrowth data set and looking at the structure of the dataframe.
    ```{r}
    library(datasets)
    data("ToothGrowth")
    str(ToothGrowth)
    ```
2. Data Summary
    First, looking at summary statistics.
    ```{r}
    summary(ToothGrowth)
    ```
    Then, looking at overall distribution of tooth growth (len), overlaying a normal distribution.
    ```{r}
    library(ggplot2)
    ggplot(data=ToothGrowth,aes(x=len))+
            geom_histogram(aes(y=..density..,fill=..count..),
                           binwidth=1,
                           color="black")+
            scale_fill_gradient("Count",
                                low="#DCDCDC",
                                high="#7C7C7C")+
            labs(title="Actual Distribution Growth vs Normal Distribution",
                 x="Growth",
                 y="Density")+
            geom_line(aes(y=dnorm(x=ToothGrowth$len,mean=mean(ToothGrowth$len),sd=sd(ToothGrowth$len))),
                      color="magenta",
                      size=1)+
            theme_classic()
    ```
    It appears that the toothgrowth sample is not really normally distributed. While there appear to be fewer observations on the low/high end, there seem to be some clustering of observations around 10, 14 and 26.
3. Comparing Tooth Growth by supplement & dosage
    a) Performing a two-sample t-test to see if the is a statistically significant difference in mean tooth growth by supp.
    H0: There is no difference in tooth growth by supplement
    Ha: There is a difference in tooth growth by supplement
    ```{r}
    with(ToothGrowth,t.test(len~supp,alternative="two.sided",mu=0,conf.level=0.95, paired=FALSE))
    ```
    Based on the t-test's p-value, which is greater than 0.05, and the 95% confidence interval around the mean difference, which includes 0, we fail to reject the H0, stating that there is no difference in toothgrowth based on supplement.
    Since there appears to be no significant difference between the supplements' effect on toothgrowth, it may be worth to investigate if there are differences in effects on toothgrowth based on dosages of the supplements.
    b) Performing two sample t-tests by dosage between the 2 supplements
        First, creating subsets of the data based on dosage levels 0.5, 1.0 and 2.0.
    ```{r}
    halfdose<-ToothGrowth[ToothGrowth$dose==0.5,]
    fulldose<-ToothGrowth[ToothGrowth$dose==1.0,]
    doubledose<-ToothGrowth[ToothGrowth$dose==2.0,]
    ```
    Next, performing three 2-sample t-tests with following hypotheses:
    H0: There is no difference in tooth growth between supplements at each of the 0.5mg/1.0mg/2.0mg levels.
    Ha: There is a difference in tooth growth between supplements at the difference dosage levels.
    
    Test for the 0.5mg dose:
    ```{r}
    with(halfdose,t.test(len~supp,alternative="two.sided",mu=0,conf.level=0.95,paired=FALSE))
    ```
    Based on the p-value being <0.05  and both ends of the confidence interval being above zero, we reject the H0 and conclude that there is a statistically significant difference in tooth growth between supplements at dose of 0.5mg.
    Test for the 1.0mg dose:
    ```{r}
    with(fulldose,t.test(len~supp,alternative="two.sided",mu=0,conf.level=0.95,paired=FALSE))
    ```
    Based on the p-value being <0.05  and both ends of the confidence interval being above zero, we reject the H0 and conclude that there is a statistically significant difference in tooth growth between supplements at dose of 1mg.
    ```{r}
    with(doubledose,t.test(len~supp,alternative="two.sided",mu=0,conf.level=0.95,paired=FALSE))
    ```
    Based on the p-value being >0.05  and the confidence interval including zero, we failt to reject the H0, stating that there is no difference in tooth growth between supplements at dose of 2mg.
4. Assumptions & Conclusions
    Assuming that these samples were not paired, at dosage levels of 0.5mg and 1mg per day, the type of supplement appears to play a role in the effect on tooth length in guinea pigs.  
    These data do not include a control group which was not receiving any supplements. So, we cannot say that these supplements increase toothgrowth beyond what would naturally occur. This analysis merely shows that at lower dosage levels (0.5mg and 1mg per day) there appears to be a difference in tooth growth while at 2mg per day there appears to be no difference. 
    When ignoring dosage levels, there appear to be no statistically significant differences between the effects of orange juice and ascorbic acid on tooth growth in guinea pigs.