---
title: "Statistical Inference Course Project"
author: "gspinnadauhu"
date: "March 23, 2017"
output: pdf_document
---

#### Part 1: Simulation Exercise
1. Overview
In this project we are investigating the exponential distribution in R and compare it with the Central Limit Theorem. We are investigating distribution of averages for 40 expenonentials.
2. Simulations
Taking 1000 averages and variances of 40 random exponential distributions with lambda=0.2. Putting samples in dataframe sims.
```{r}
lambda<-0.2
mns<-NULL
vars<-NULL
for (i in 1:1000) {
        mns=c(mns,mean(rexp(40,lambda)))
        vars=c(vars,var(rexp(40,lambda)))
}
sims<-data.frame(mns)
```
3. Sample Mean vs Theoretical Mean
The theoretical mean of an exponential distribution is 1/lambda (lambda=0.2 in this case).
Comparing the theoretical mean with the mean of sampled means.
```{r}
library(ggplot2)
ggplot(data=sims, aes(mns))+
        geom_histogram()+
        labs(title="Theoretical Mean vs Sample Mean",
             x="Sample Means",
             y="Count")+
        geom_vline(xintercept=1/lambda, colour="green",linetype="longdash", size=1)+
        geom_text(aes(x=1/lambda*1.15,y=10,label="Theoretical Mean"),color="green")+
        theme_classic()
```
As visible above, the sample means appear to center around the distribution's theoretical mean.
Furthermore, the mean of the sample means is almost exactly equal to 1/lambda=1/0.2=5
```{r}
mean(sims$mns)
```
4. Sample Variance vs Theoretical Variance
The theoretical variance of an exponential distribution is (1/lambda)^2. Comparing the theoretical variance with the variance of the sample means.
```{r}
library(ggplot2)
ggplot(data=sims, aes(vars))+
        geom_histogram()+
        labs(title="Theoretical Variance vs Sample Variance",
             x="Sample Variances",
             y="Count")+
        geom_vline(xintercept=(1/lambda)^2, colour="blue",linetype="longdash", size=1)+
        geom_text(aes(x=(1/lambda)^2*1.6,y=10,label="Theoretical Variance"),color="blue")+
        theme_classic()
```
As visible above, the sample variances appear are for the most part located around the theoretical variance.
Furthermore, the mean of the sample variances is very close to (1/lambda)^2=(1/0.2)^2=25
```{r}
mean(sims$vars)
```


5. Distribution
Comparing 1000 draws from an exponential distribution with lambda=0.2 with the distribution of the 1000 sampled means of 40 exponentials in mns.
First, simulating the 1000 draws, converting to a data frame & plotting the distribution.
```{r}
draws<-rexp(1000,0.2)
draws<-data.frame(draws)
drawmean<-mean(draws$draws)
library(ggplot2)
ggplot(data=draws, aes(draws))+
        geom_histogram()+
        labs(title="1000 Random Draws from Exponential Distribution",
             x="Samples",
             y="Count")+
        geom_vline(xintercept=mean(drawmean), colour="lightblue",linetype="longdash", size=1)+
        geom_text(aes(x=drawmean+1.5,y=10,label="Mean"),color="lightblue")+
        theme_classic()
```
These random draws conform to the expected shape of an exponential distribution.

Now, comparing this to the distrbution of means drawn from 1000 simulated sets of 40 exponential distributions in sims$mns.
```{r}
library(ggplot2)
ggplot(data=sims, aes(mns))+
        geom_histogram()+
        labs(title="Theoretical Mean vs Sample Mean",
             x="Sample Means",
             y="Count")+
        geom_vline(xintercept=1/lambda, colour="green",linetype="longdash", size=1)+
        geom_text(aes(x=mean(sims$mns)+.25,y=10,label="Mean"),color="green")+
        theme_classic()
```
While both have a mean of 5, which matches the theoretical mean, the draws are normally distributed only in the set of 1000 means drawn for samples size 40. As the sample size increases, the sampled means conform closer to the theoretical mean, which is why the distribution looks more bell curved.

#### Part 2: Basic Inferential Data Analysis
1. Loading Data & Basic Analysis
First, loading the ToothGrowth data set and looking at the structure of the dataframe.
```{r}
library(datasets)
data("ToothGrowth")
str(ToothGrowth)
```
2. Data Summary
First, looking at summary statistics.
```{r}
summary(ToothGrowth)
```
Then, looking at overall distribution of tooth growth (len), overlaying a normal distribution.
```{r}
library(ggplot2)
ggplot(data=ToothGrowth,aes(x=len))+
        geom_histogram(aes(y=..density..,fill=..count..),
                       binwidth=1,
                       color="black")+
        scale_fill_gradient("Count",
                            low="#DCDCDC",
                            high="#7C7C7C")+
        labs(title="Actual Distribution Growth vs Normal Distribution",
             x="Growth",
             y="Density")+
        geom_line(aes(y=dnorm(x=ToothGrowth$len,mean=mean(ToothGrowth$len),sd=sd(ToothGrowth$len))),
                  color="magenta",
                  size=1.5)+
        theme_classic()
```
It appears that the toothgrowth sample is not really normally distributed. While there appear to be fewer observations on the low/high end, there seem to be some clustering of observations around 10, 14 and 26.
3. Comparing Tooth Growth by supp & dose

4. Assumptions & Conclusions