---
title: "Statistical Inference Course Project"
author: "gspinnadauhu"
date: "March 23, 2017"
output: pdf_document
---

#### Part 1: Simulation Exercise
1. Overview
In this project we are investigating the exponential distribution in R and compare it with the Central Limit Theorem. We are investigating distribution of averages for 40 expenonentials.
2. Simulations
Taking 1000 averages and variances of 40 random exponential distributions with lambda=0.2. Putting samples in dataframe sims.
```{r}
lambda<-0.2
mns<-NULL
vars<-NULL
for (i in 1:1000) {
        mns=c(mns,mean(rexp(40,lambda)))
        vars=c(vars,var(rexp(40,lambda)))
}
sims<-data.frame(mns)
```
3. Sample Mean vs Theoretical Mean
The theoretical mean of an exponential distribution is 1/lambda (lambda=0.2 in this case).
Comparing the theoretical mean with the mean of sampled means.
```{r}
library(ggplot2)
ggplot(data=sims, aes(mns))+
        geom_histogram()+
        labs(title="Theoretical Mean vs Sample Mean",
             x="Sample Means",
             y="Count")+
        geom_vline(xintercept=1/lambda, colour="green",linetype="longdash", size=1)+
        geom_text(aes(x=1/lambda*1.15,y=10,label="Theoretical Mean"),color="green")+
        theme_classic()
```
As visible above, the sample means appear to center around the distribution's theoretical mean.
Furthermore, the mean of the sample means is almost exactly equal to 1/lambda=1/0.2=5
```{r}
mean(sims$mns)
```
4. Sample Variance vs Theoretical Variance
The theoretical variance of an exponential distribution is (1/lambda)^2. Comparing the theoretical variance with the variance of the sample means.
```{r}
library(ggplot2)
ggplot(data=sims, aes(vars))+
        geom_histogram()+
        labs(title="Theoretical Variance vs Sample Variance",
             x="Sample Variances",
             y="Count")+
        geom_vline(xintercept=(1/lambda)^2, colour="blue",linetype="longdash", size=1)+
        geom_text(aes(x=(1/lambda)^2*1.6,y=10,label="Theoretical Variance"),color="blue")+
        theme_classic()
```
As visible above, the sample variances appear are for the most part located around the theoretical variance.
Furthermore, the mean of the sample variances is very close to (1/lambda)^2=(1/0.2)^2=25
```{r}
mean(sims$vars)
```


5. Distribution
Comparing 1000 draws from an exponential distribution with lambda=0.2 with the distribution of the sampled means in mns.
```{r}

```

#### Part 2: Basic Inferential Data Analysis
1. Loading Data & Basic Analysis
2. Data Summary
3. Comparing Tooth Growth by supp & dose
4. Assumptions & Conclusions